The Bolocam Galactic Plane Survey: Survey Description and Data Reduction

### 2009-04-16 methods/ms.tex
new tests for polynomial subtraction to try to get rid of bowls

% Figure 3
\Figure{bolocam_bandpass}{The Bolocam 1.1 mm bandpass.  Bright
molecular emission lines are shown at their rest frequencies.  Note
that the Bolocam passband rejects $>90\%$ of the $^{12}$CO flux,
leaving our continuum measurements largely uncontaminated.  SO and
CH$_3$OH lines are probably the dominant contributors to line flux in
the passband.  \citet{nummelin1998} found that 22\% of the flux in one
pointing towards Sgr B2 was from lines, and \citet{yoshida2005}
reported $>40\%$ in Orion A from line emission over the 260-328 GHz band. \TBD\ Other lines
lying in the Bolocam passband include CS($5\to4$) and ($6\to5$) (245
and 293 GHz), HCN($3\to2$) (265 GHz) and HCO$^+(3\to2)$ (267 GHz).}
{fig:Bandpass}{1.0}

@@

We have convolved the fits files\footnote{{\tt
ftp://cdsarc.u-strasbg.fr/pub/cats/J/A+A/476/1243/fits/}} with a
Gaussian kernel that generates an approximately 35\arcsec\ beam,
matched to our images.  The results are shown in Figure
\ref{fig:CygnusComparison}.

\Table
{llllll}
{Comparison of \citet{motte07} IRAM measurements to BGPS measurements
in the Cygnus X region
%Note:  The \citet{motte07} integration box is given in terms of pc. 
%The assumed distance must be about 1.7 kpc \citep{schneider2006}    % no, it's 1.7kpc stated explicitly
}
{
%Source  &   BGPS           &  Motte \\
%        &   peak mJy/beam  &  mJy in box (see Motte+ Table 1)
Motte Source Number &  Motte Table 1 integrated flux &  Motte scaled Jy/bm  & BGPS Jy/Bm &    Comments \\
}
{tab:CygnusComparison}
{
3  &    1.59 [0.11x0.09] & 2.61      &  2.30  &  DR 17                               \\
11 &    0.34 [0.11x0.06] & 0.48      &  0.50  &  West SW of DR21                     \\
12 &    1.63 [0.11x0.09] & 2.60      &  2.20  &    "                                 \\
16 &    0.55 [0.12x0.12] & 0.82      &  0.72  &    "                                 \\
26 &    0.34 [0.20x0.10] & 0.48      &  0.44  &    "                                 \\
29 &    0.54 [0.25x0.13] & 1.10      &  0.62  &    "                                 \\
30 &   10.54 [0.15x0.10] &13.5       &  9.80  &  W75(N)                              \\
44 &    8.43 [0.14x0.11] &14.5       & 12.7   &  DR21 (OH)                           \\
46 &   17.90 [0.08x0.03] &18.4       & 15.8   &  DR21 f-f, maser (H2O)               \\
48 &    3.72 [0.17x0.11] & 9.4       &  7.8   &  DR21 (OH)-S                         \\
51 &    2.37 [0.24x0.03] & 5.5       &  4.1   &  DR21 FIR2                           \\
53 &    1.60 [0.14x0.09] & 4.3       &  2.7   &  North of DR21  FIR3                 \\
56 &    0.60 [0.12x0.08] & 0.81      &  0.75  &  SE of DR21 47                       \\
58 &    2.73 [0.23x0.19] & 2.40      &  1.64  &  IRAS 20375+4109                     \\
59 &    0.37 [0.12x0.05] & 1.3       &  0.93  &  East of IRAS 20375+4109             \\
60 &    0.56 [0.16x0.13] & 1.00      &  0.78  &  East of IRAS 20375+4109             \\
61 &    0.23 [0.12x0.09] & 0.60      &  0.49  &  MSX 81.5168+0.01926                 \\
62 &    0.50 [0.18x0.12] & 1.23      &  0.80  &  between 63 \& IRAS 20375+4109       \\
63 &    1.09 [0.08x0.05] & 1.60      &  1.50  &  DR22                                \\
65 &    0.78 [0.16x0.12] & 1.15      &  1.10  &                                      \\
66 &    0.25 [0.15x0.06] & 0.63      &  0.76  &                                      \\
68 &    0.84 [0.10x0.10] & 1.01      &  0.94  &                                      \\
69 &    1.28 [0.47x0.18] & 1.12      &  1.20  &                                      \\
72 &    0.54 [0.28x0.13] & 0.58      &  0.47  &                                      \\
}

\FigureTwo{fluxcal_errorbars}{calib_vs_dc}
%{rel_flux_cal_050303_ob3_to_050405_o66}
%{calibration_plot_2004}
{Left: Average calibration curve [V/Jy] versus the mean detector voltage, a
proxy for atmospheric loading.  Red points are observations of Mars and black
of Uranus.  The black line is a 2nd-order fit with 0,0 (no response if no
measurable potential difference) forced.  Right: Scaling of the relative
response to the atmosphere for a single detector compared to the array
median.}{fig:CalibrationCurves}{1.0}

@@

Pointing observations were performed approximately once every two
hours over the night.  These sources were then mapped in alt/az
coordinates relative to the nominal telescope boresight.  The
elevation offsets showed a deviation from zero which was empirically
well-modeled by a quadratic function of elevation and a linear
function of azimuth.  No systematic deviation was observed for the
azimuth offsets.

Because the Galactic Plane appears over much of the sky, the pointing
model necessarily encompasses a large range of elevation and azimuth.
Nevertheless, an RMS scatter of $\sim 6\arcsec$ for the model over the
entire observed range was achieved, with the scatter somewhat worse at
the highest ZA.  A separate model was constructed for each epoch.  The
residuals are nearly Gaussian, with an RMS of $\sim6\arcsec$.

@@

In addition to the model for the pointing center, it is necessary to
empirically determine actual projected pattern of the array on the
sky, and measure the offset angle between the focal plane and the sky
coordinate system.  This is done by making observations which track
all detectors across a bright source (a planet) and making maps from each
bolometer individually.

@@

A mean focus offset for the secondary is determined once per observing
run.  An elevation-dependent correction is applied via a
CSO-determined look-up table between raster scans.  We do not see any
evidence that more frequent focus checks improve the quality of the
PSF.

@@

This method of alignment avoids the ambiguities inherent in using
extracted sources to align fields, as the Bolocam sources are rarely
point-like.  It further avoids the pointing smearing and loss of peak
flux density which would result if the the maps were coadded using the
pointing model alone, where each individual observation would be
coadded with tthe 6\arcsec\ RMS uncertainty of the model.  The
relative alignment using the correlation maps is typically accurate to
\TBD\ $<1\arcsec$.

TBD: In low s/n fields, specifically \lon=65 to 75, there was not enough signal to
acquire a pointing offset using cross-correlation.  In these fields the
relative pointing offset is larger and the beam is slightly larger.

@@

All of the SCUBA 850 and 450 \mum\ data has been re-processed in a
uniform manner and made publicly available as the SCUBA Legacy Catalog
(hereafter SLC).  These maps provide an excellent cross-check on the
accuracy of the BGPS pointing model.  We have used the same
cross-correlation procedure as that used to obtain the relative
alignment of Bolocam observations to compare to the SLC.  A number of
concerns might arise in comparing the Bolocam to the SCUBA maps,
including the differing scan strategies (SCUBA maps were typically
taken in jiggle-map mode, which tends to remove extended structure)
and the often restricted range of positions measured within an
extended cloud for the SLC.  The cross-correlation technique actually
mitigates these effects, since it does not require the establishment
of a single position for an object (which may differ due to the change
in wavlength between the surveys or due to the observing strategy) but
rather makes use of all the information available.  In the event, the
morphological correlation between SCUBA and BGPS sources is generally
excellent, giving a high degree of confidence to the comparison and to
the results obtained from the pointing model.  An example is shown in
Figure \ref{fig:SCUBAPointingComparison}.  We find that the RMS offset
between BGPS and SLC positions within a $3\arcdeg \times 1\arcdeg$
bloc defined by the BGPS is \TBD\ a few arcseconds, consistent with
the errors derived for the Bolocam pointing model.

>>>>

% Figure 3
\Figure{bolocam_bandpass}{The Bolocam 1.1 mm bandpass.  Bright
molecular emission lines are shown at their rest frequencies.  Note
that the Bolocam passband rejects $>90\%$ of the $^{12}$CO flux,
leaving our continuum measurements largely uncontaminated.  SO and
CH$_3$OH lines are probably the dominant contributors to line flux in
the passband.  \citet{nummelin1998} found that 22\% of the flux density in one
pointing towards Sgr B2 was from lines, and \citet{yoshida2005}
reported $>40\%$ in Orion A from line emission over the 260-328 GHz band. \TBD\ Other lines
lying in the Bolocam passband include CS($5\to4$) and ($6\to5$) (245
and 293 GHz), HCN($3\to2$) (265 GHz) and HCO$^+(3\to2)$ (267 GHz).}
{fig:Bandpass}{1.0}

@@

We have convolved the \citet{motte07} fits files\footnote{{\tt
ftp://cdsarc.u-strasbg.fr/pub/cats/J/A+A/476/1243/fits/}} with a Gaussian
kernel that generates an approximately 35\arcsec\ beam, and compared the
measured fluxes with those from the BGPS.  The results are shown in Figure
\ref{fig:CygnusComparison}.

\Table
{llllll}
{Comparison of \citet{motte07} IRAM measurements to BGPS measurements
in the Cygnus X region.  The IRAM data has been convolved to the BGPS beam
to make the flux density measurements.
%Note:  The \citet{motte07} integration box is given in terms of pc. 
%The assumed distance must be about 1.7 kpc \citep{schneider2006}    % no, it's 1.7kpc stated explicitly
}
{
%Source  &   BGPS           &  Motte \\
%        &   peak mJy/beam  &  mJy in box (see Motte+ Table 1)
Motte Source &  Motte Table 1   &  Motte scaled  & BGPS  &  Comments \\
      Number &  integrated flux &  Jy/bm         & Jy/bm &           \\
}
{tab:CygnusComparison}
{
3  &    1.59 [0.11x0.09] & 2.61      &  2.30  &  DR 17                               \\
11 &    0.34 [0.11x0.06] & 0.48      &  0.50  &  West SW of DR21                     \\
12 &    1.63 [0.11x0.09] & 2.60      &  2.20  &    "                                 \\
16 &    0.55 [0.12x0.12] & 0.82      &  0.72  &    "                                 \\
26 &    0.34 [0.20x0.10] & 0.48      &  0.44  &    "                                 \\
29 &    0.54 [0.25x0.13] & 1.10      &  0.62  &    "                                 \\
30 &   10.54 [0.15x0.10] &13.5       &  9.80  &  W75(N)                              \\
44 &    8.43 [0.14x0.11] &14.5       & 12.7   &  DR21 (OH)                           \\
46 &   17.90 [0.08x0.03] &18.4       & 15.8   &  DR21 f-f, maser (H2O)               \\
48 &    3.72 [0.17x0.11] & 9.4       &  7.8   &  DR21 (OH)-S                         \\
51 &    2.37 [0.24x0.03] & 5.5       &  4.1   &  DR21 FIR2                           \\
53 &    1.60 [0.14x0.09] & 4.3       &  2.7   &  North of DR21  FIR3                 \\
56 &    0.60 [0.12x0.08] & 0.81      &  0.75  &  SE of DR21 47                       \\
58 &    2.73 [0.23x0.19] & 2.40      &  1.64  &  IRAS 20375+4109                     \\
59 &    0.37 [0.12x0.05] & 1.3       &  0.93  &  East of IRAS 20375+4109             \\
60 &    0.56 [0.16x0.13] & 1.00      &  0.78  &  East of IRAS 20375+4109             \\
61 &    0.23 [0.12x0.09] & 0.60      &  0.49  &  MSX 81.5168+0.01926                 \\
62 &    0.50 [0.18x0.12] & 1.23      &  0.80  &  between 63 \& IRAS 20375+4109       \\
63 &    1.09 [0.08x0.05] & 1.60      &  1.50  &  DR22                                \\
65 &    0.78 [0.16x0.12] & 1.15      &  1.10  &                                      \\
66 &    0.25 [0.15x0.06] & 0.63      &  0.76  &                                      \\
68 &    0.84 [0.10x0.10] & 1.01      &  0.94  &                                      \\
69 &    1.28 [0.47x0.18] & 1.12      &  1.20  &                                      \\
72 &    0.54 [0.28x0.13] & 0.58      &  0.47  &                                      \\
}

\FigureTwo{fluxcal_errorbars}{calib_vs_dc}
%{rel_flux_cal_050303_ob3_to_050405_o66}
%{calibration_plot_2004}
{Left: Average calibration curve [V/Jy] versus the mean detector voltage, a
proxy for atmospheric loading.  Red points are observations of Mars and black
of Uranus.  The black line is a 2nd-order fit with 0,0 forced (no response if no
measurable potential difference) .  Right: Scaling of the relative
response to the atmosphere for a single detector compared to the array
median.}{fig:CalibrationCurves}{1.0}

@@

Pointing observations were performed approximately once every two
hours over the night.   The
elevation offsets showed a deviation from zero which was empirically
well-modeled by a quadratic function of elevation and a linear
function of azimuth.  No systematic deviation was observed for the
azimuth offsets.

Because the Galactic Plane appears over much of the sky, the pointing
model necessarily encompasses a large range of elevation and azimuth.
Nevertheless, an RMS scatter of $\sim 6\arcsec$ for the model over the
entire observed range was achieved, with the scatter somewhat worse at the
highest ZA.  A separate model was constructed for each epoch, but epoch V was
used as the primary pointing calibration and is shown in figure
\ref{fig:PointingModel}.

@@

In addition to the model for the pointing center, it is necessary to
empirically determine actual projected pattern of the array on the
sky, and measure the offset angle between the focal plane and the sky
coordinate system.  This is done by making observations which track
all detectors across a bright source (e.g. a planet) and making maps from each
bolometer individually.

@@

A mean focus offset for the secondary is determined once per observing
run.  An elevation-dependent correction is applied via a
CSO-determined look-up table between raster scans.  We did not see any
evidence that more frequent focus checks improve the quality of the
PSF.

@@

This method of alignment avoids the ambiguities inherent in using
extracted sources to align fields, as the Bolocam sources are rarely
point-like.  It further avoids the pointing smearing and loss of peak
flux density which would result if the the maps were coadded using the
pointing model alone, where each individual observation would be
coadded with tthe 6\arcsec\ RMS uncertainty of the model.  The
relative alignment using the correlation maps is typically accurate to
 $<1\arcsec$.

In low s/n fields, specifically \lon=65 to 75, there was not enough signal to
acquire a pointing offset using cross-correlation.  In these fields the
relative pointing offset is larger and the beam is slightly larger.

@@

All of the SCUBA 850 and 450 \mum\ data has been re-processed in a
uniform manner and made publicly available as the SCUBA Legacy Catalog
\citep[][hereafter SLC]{difrancesco08}.  These maps provide an excellent cross-check on the
accuracy of the BGPS pointing model.  We have used the same
cross-correlation procedure as that used to obtain the relative
alignment of Bolocam observations to compare to the SLC.  A number of
concerns might arise in comparing the Bolocam to the SCUBA maps,
including the differing scan strategies (SCUBA maps were typically
taken in jiggle-map mode, which tends to remove extended structure)
and the often restricted range of positions measured within an
extended cloud for the SLC.  The cross-correlation technique actually
mitigates these effects, since it does not require the establishment
of a single position for an object (which may differ due to the change
in wavlength between the surveys or due to the observing strategy) but
rather makes use of all the information available.  The
morphological correlation between SCUBA and BGPS sources is generally
excellent, giving a high degree of confidence to the comparison and to
the results obtained from the pointing model.  An example is shown in
Figure \ref{fig:SCUBAPointingComparison}.  We find that the RMS offset
between BGPS and SLC positions within a $3\arcdeg \times 1\arcdeg$
bloc defined by the BGPS is \TBD\ a few arcseconds, consistent with
the errors derived for the Bolocam pointing model.
